---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", cache = TRUE)

```

# quarrisk

<!-- badges: start -->
<!-- badges: end -->

`quarrisk` is an R package that helps you understand and visualise the risk of quarantine breaches under different COVID-19 quarantine and vaccination scenarios.
It turns simulation outputs into clear summaries and includes a simple interactive Shiny dashboard.

## What this package does

- Provides a built-in dataset (breach_data) that summarises simulated infection outcomes for travellers and workers.

- Lets users compare breach risk across scenarios (e.g., vaccinated vs unvaccinated, home vs hotel).

- Includes a Shiny app that shows interactive time-series plots.

- Documents every function and dataset with roxygen2 help pages and a vignette.

## Installation

You can install the development version of quarrisk from [GitHub](https://github.com/) with:

```{r install, eval=FALSE}
# Development install (do NOT run during README knit)
# install.packages("pak")
pak::pak("ETC5523-2025/assignment-4-packages-and-shiny-apps-ft-siddhi")
```

```{r, eval=FALSE}
install.packages("remotes")
remotes::install_github("ETC5523-2025/assignment-4-packages-and-shiny-apps-ft-siddhi")
```


### Load the package and view the data

```{r, load-breach-data, message=FALSE}
library(quarrisk)
data("breach_data", package = "quarrisk")
stopifnot(exists("breach_data"))
# quick sanity peek
utils::head(breach_data)
```

### Make a small summary and plot

```{r, plot-daily, message=FALSE, warning=FALSE, cache=TRUE}

set.seed(1)
small <- dplyr::slice_sample(breach_data, n = min(5000, nrow(breach_data)))

daily <- small |>
  dplyr::filter(!is.na(.data$integrated_FoI)) |>
  dplyr::mutate(day = floor(.data$days_infectious_community)) |>
  dplyr::summarise(
    mean_integrated_FoI = mean(.data$integrated_FoI, na.rm = TRUE),
    .by = c(scenario, day)
  )

ggplot2::ggplot(daily, ggplot2::aes(day, mean_integrated_FoI, colour = scenario)) +
  ggplot2::geom_line(linewidth = 1) +
  ggplot2::labs(
    title = "Average Integrated Force of Infection by Scenario",
    x = "Days infectious in community",
    y = "Average integrated FoI",
    colour = "Scenario"
  ) +
  ggplot2::theme_minimal()

```

### Launch the interactive app

```{r, eval=FALSE}
quarrisk::run_quarrisk()
```


### Vignette

For a guided explanation (background, methods, and visual interpretation):

```{r, eval=FALSE}
browseVignettes("quarrisk")
```


### License

MIT License — see the `LICENSE` file for full text.

### Source acknowledgement

The dataset was derived from the open-source COVIDQuarantine modelling repository (Mike Lydeamore et al., 2021) and simplified for teaching purposes.
All credit to the original authors for their modelling framework.

### Summary

`quarrisk` demonstrates how simulation data can be wrapped into a tidy, documented R package with a reproducible workflow and an interactive dashboard making epidemiological insights accessible to anyone.

### Documentation website

[![pkgdown docs](https://img.shields.io/badge/docs-pkgdown-blue.svg)](https://etc5523-2025.github.io/assignment-4-packages-and-shiny-apps-ft-siddhi/)
